%\documentclass[oribibl]{llncs}
\documentclass[a4paper]{report}
\usepackage{wrapfig}
\usepackage{times}
\usepackage{stmaryrd, latexsym, amsmath, amssymb, wasysym}
\usepackage{fancyvrb}
\usepackage[Lenny]{fncychap}
\usepackage{listings}
\usepackage{url}
\usepackage{semantic}
\usepackage{caption}
\usepackage{array}
\usepackage{graphicx}
\newcommand{\nametklass}{taint-aware }
\newcommand{\nameTklass}{Taint-aware }
\newcommand{\namefunc}{taint-aware }
\newcommand{\suggestions}{Suggestions for lunch \texttrademark }

\newcommand{\arrowop}[1]{$#1\negthickspace #1\negthickspace #1$}
\newcommand{\co}[1]{$\cod{#1}$}
% security type with index
\newcommand{\sts}[1]{s_{#1}^l}
% security type without index
\newcommand{\st}{s^l}
% symbol of constraint IS
\newcommand{\is}{\sim}
% symbol of gconstraint
\newcommand{\guard}{\lhd}
\newcommand{\sleql}{\LHD}
\newcommand{\lleqs}{\preceq}
% tag constraint
\newcommand{\tagup}{\uparrow}
% declassification constraint
\newcommand{\decl}{\downarrow}
% lifting function
\newcommand{\lift}{\nearrow}
% basic type without index
\newcommand{\typ}{\tau}
% basic type with index
\newcommand{\typn}[1]{\tau_{#1}}
% result of type system
\newcommand{\res}[2]{{#1}\mid {#2}}


%\newenvironment{code}{\begin{Verbatim}[fontsize=\footnotesize]}{\end{Verbatim}}

\pagestyle{plain}

\title{A library for Taint Analysis in Python: discussions and implementation}

\newcommand{\myTitle}{A library for Taint Analysis in Python: discussions and implementation}
\newcommand{\mySubtitle}{}
\newcommand{\myAuthor}{Ing. Juan Jos\'{e} Conti}
%\institute{
%Department of Computer Science,
%Chalmers University of Technology\\
%412 96 G\"{o}teborg, Sweden
%}

\author{Juan Jos\'{e} Conti}

\begin{document}

\pagestyle{plain}
\pagenumbering{roman}

%\maketitle
\thispagestyle{empty}

\clearpage
\par\vskip 2cm
\begin{center}
{\Huge\bf \myTitle                                         %PERSONALIZE
\vskip 1cm \Large \mySubtitle                              }           %PERSONALIZE
\par\vspace {6cm}
{\large Graduate Thesis }
\par\vspace {1cm}
{\large ????????????????????\\
%at \\
UTN  }
\par \vspace{1cm}
{\large September 2010  }            %PERSONALIZE
\par\vspace {1cm} {\large by}
\par \vspace {1cm}
{\Large \myAuthor            }                                  %PERSONALIZE
\par\vspace {1cm}
{\large Santa Fe, Argentina}                 %PERSONALIZE
\end{center}

\clearpage

\thispagestyle{empty}
\noindent
Supervisor: Dr. Alejandro Russo \\
Department of Computer Science and Engineering \\
Chalmers University of Technology \\
SE - 412 96 G\"{o}teborg, Sweden \\

\noindent
Co-supervisor: Ms. Susana Romaniz\\
Facultad Regional Santa Fe \\
Universidad Tecnol\'{o}gica Nacional \\
Argentina\\

\vfill\noindent
Copyright \copyright\ 2010 by \myAuthor \\
E-mail: jjconti@frsf.utn.edu.ar

\clearpage

\tableofcontents

\listoffigures
\addcontentsline{toc}{chapter}{List of Figures}

\listoftables
\addcontentsline{toc}{chapter}{List of Tables}

\chapter*{Abstract}
\addcontentsline{toc}{chapter}{Abstract}


Vulnerabilities in web applications present threats to on-line systems.
SQL injection and cross-site scripting attacks are among the
most common threats found nowadays. These attacks are
often result of improper or none input validation. 
To help discover such vulnerabilities, 
popular web scripting 
languages like Perl, Ruby, PHP, 
and Python perform taint analysis.
The analysis is usually implemented using an static (e.g. type
systems) or dynamic (e.g. execution monitors) techniques. 
In the latter case, the Perl, Ruby, PHP, and Python interpreters have
been adapted to provide a taint mode. 
However, modifying interpreters might be a major task in its own
right. In fact, it is very probably that  
new releases of interpreters require to 
be adapted to provide a taint mode.
Differently from previous approaches, 
we show how to provide taint analysis for Python via a library
 written entirely in Python, and thus avoiding modifications in the interpreter.
The concepts of classes, decorators and dynamic dispatch
makes our solution lightweight, easy to use, and particularly neat.
With minimal or none effort, the library can be adapted
to work with different Python interpreters.



\chapter*{Acknowledgement}
\addcontentsline{toc}{chapter}{Acknowledgments}

This work wouldn't be possible without kindly support from many people.

First of all, I would like to express my immense gratitude to my dear supervisor, 
Dr. John Hughes, who has suggested me such an exciting topic, has provided me brilliant ideas
when I met difficulties, and has given me insightful advice as well as zealous encouragement.
I cannot overstate my grateful to Alejandro Russo. This work wouldn't look like it is today
without his clever ideas, enthusiastic discussions, and helpful comments.
Every meeting with them is informative and enjoyable for me.  
I learn a lot not only from their professional expertise but also from their optimistic personalities
when facing problems.
Working with them is definitely an unforgettable experience in my life.

I would like to thank Dr. David Sands, Dr. Andrei Sabelfeld, Aslan Askarov, and people in the Multi
Group at Department of Computer Science and Engineering for providing useful comments on many 
parts of this work. I also like to thank B\"{o}rje Johansson for helping me facilitate the 
whole process.

I want to dedicate this work to my dearest family: my grand mother, Cho-she Tsai, who always fully supports
me since I was a child, my father, Cheng-chih Tsai, and my mother, Li-ching Chao, who encourage me
to pursue my dream and support me with infinite love, and my brother, Yu-chung Tsai, with whom I learn, play, 
and grow. I am also grateful to my lovely relatives whose constantly care and greetings make me
feel warm even in the cold winter of Sweden.

I cannot continue without thanking a group of extraordinary friends. They are Li-ting Chao, 
Chien-chih Chen, Chien-huei Chen, Fei-neng Chuang, Yu-ting Liao, Yueh-ting Liao as well as
other members of NCTU Europe. We share the happiness and the bitterness of studying in Sweden and
strive for graduation together. I will miss our wonderful weekly parties and movie nights.

In the end, I like to express my gratitude to all my friends in Taiwan for helping me
in every aspect and making me feel connected despite far from home.
In particular, I would like to thank Cheng-dar Li, Yi-tzu Lin, and Fu-kuo Tseng.


{\small{
\paragraph{Acknowledgments}
Thanks are due to Arnar Birgisson for interesting discussions.
This work was funded by the Swedish research agencies VR and 
the scholarship program for graduated students 
from the Universidad Tecnol\'{o}gica Nacional, Facultad Regional Santa Fe. 
}}

% real report begins
\cleardoublepage
\pagestyle{plain}
\pagenumbering{arabic}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% 1 INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Introduction}
\label{Chap:Introduction}
%\subimport{01-Introduction}{intro.tex}
\input{01-introduction.tex}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% 2 Taint Analysis %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Taint Analysis in general + related works}
\label{Chap:taint}
\input{02-taint.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% 3 A library for Taint Analysis in Python %%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{A library for Taint Analysis in Python}
\label{Chap:Library}
\input{03-library.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% 4  Implementation %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Implementation}
\label{Chap:Implementation}
\input{04-implementation.tex}
        generation of classe
        taint aware
        propagate func
        propagate method

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% 5  Limitations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Limitations}
\label{Chap:Limitations}
\input{05-limitations.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% 6  Conclusions %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{Conclusions and future works}
\label{Chap:Conclusions}
\input{06-conclusion.tex}

\section{Future Work}


\section{Conclusion}
% Function : sequential information flow, references, precise security types, 
%            less burden of programmers(unification), close internal timing channel,
%            case study
% Advantage : low cost to migrate, practical, light-weighted, scheduler independent, 
%             full permisiveness as before.
% Cons : reference checking is delayed to run-time, debug message is not clear
FlowHaskellRef extends FlowHaskell with reference manipulation and secure multithreaded programming.
The new contributions are listed as following:
\begin{enumerate}
\item Complex security types permits more accurate description of data
\item Reference manipulation provides possibility of shared resources
\item Unification mechanism infers security types automatically and mitigates the responsibility of programmers
\item Scheduler-independent run-time system eliminates {\em internal-timing} channels in multithreaded programming
\item Two full case studies to evaluate FlowHaskellRef
\end{enumerate}


\chapter*{Appendix}
\addcontentsline{toc}{chapter}{Appendix}

\setcounter{section}{0}
\renewcommand\thesection{\Alph{section}}

\section{Library source code}

\section{Test cases}

\section{\suggestions source code}



\bibliographystyle{abbrv}
\bibliography{literature}

\end{document}
